Management induced changes of soil organic carbon on global croplands

### 2022-07-10 SOCbudgetPaper/SOCbudget.Rmd
update

  Soil organic carbon (SOC) is one of the largest terrestrial carbon (C) stocks on Earth. The first meter of the Earth's soils profile stores three times as much C as the vegetation and twice the amount of C in the atmosphere. SOC has been depleted by anthropogenic land-cover change and agricultural management. However, the latter has so far not been well represented in global C stock assessments. While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale. Here we develop a spatially explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices. We combine it with a reduced-complexity model based on the IPCC Tier 2 method to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils. We estimate that due to arable farming, soils have lost around $34.6\,\unit{GtC}$ relative to a counterfactual hypothetical natural state in 1975. Within the period 1975--2010 this SOC debt continued to expand by $5\,\unit{GtC}$ ($0.14\,\unit{GtC\,yr^{-1}}$) to around 39.6 GtC. However, accounting for historical management led to $2.1\,\unit{GtC}$ less ($0.06\,\unit{GtC\,yr^{-1}}$) emissions than under the assumption of constant management. We also find that management decisions have influenced the historical SOC trajectory most strongly by residue returning, indicating that increasing SOC sequestration by biomass retention may be a promising negative emissions technique. The reduced-complexity SOC model may allow to simulate management-induced SOC sequestration also within computationally demanding integrated (land-use) assessment modeling. 

@@

>>>>

  Soil organic carbon (SOC), one of the largest terrestrial carbon (C) stocks on Earth, has been depleted by anthropogenic land-cover change and agricultural management. However, the latter has so far not been well represented in global C stock assessments. While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale. Here we develop a spatially explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices. We combine it with a reduced-complexity model based on the IPCC Tier 2 method to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils. We estimate that due to arable farming, soils have lost around $34.6\,\unit{GtC}$ relative to a counterfactual hypothetical natural state in 1975. Within the period 1975--2010 this SOC debt continued to expand by $5\,\unit{GtC}$ ($0.14\,\unit{GtC\,yr^{-1}}$) to around 39.6 GtC. However, accounting for historical management led to $2.1\,\unit{GtC}$ less ($0.06\,\unit{GtC\,yr^{-1}}$) emissions than under the assumption of constant management. We also find that management decisions have influenced the historical SOC trajectory most strongly by residue returning, indicating that increasing SOC sequestration by biomass retention may be a promising negative emissions technique. The reduced-complexity SOC model may allow to simulate management-induced SOC sequestration also within computationally demanding integrated (land-use) assessment modeling. 

@@

    ```{=latex}
  \begin{figure}[H]
    \includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scatterComparePointData_2010.png} 
    \caption{Correlation between modeled and measured $SOC$ stocks. Given the wide span between minimun and maximum measured $SOC$ stocks within in a given cell, we correlated median values with our modeled results. Both cropland ($R2=0.09$) and areas with natural vegetation ($R2=0.08$) tend to be lower in our results than in the point measurements.}
    \label{fig:SOCpoint}
  \end{figure}
   ```

     ```{=latex}
  \begin{figure}[H]
    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/scatterCompareLPJmL42010_pnv.png}
  \caption{Correlation between modeled SOC stocks of LPJmL4 and this study for an hypothetical potential natural state (PNV) for the year 2010. The grey lines indicate the 1:1 line.}
    \label{fig:SOCLPJmL4}
  \end{figure}
   ```
